\texttt{vqtl}: An \texttt{R} package for Mean-Variance QTL Mapping

### 2018-05-06 Package_vqt_Manuscript.tex
edits

It has long been recognized that the extent of residual variation is itself a heritable trait \citep{Lynch1998}.
This phenomenon has been described theoretically \citep{Hill2010,Hill2004-uo} and demonstrated in inbred model organisms \citep{Sorensen2015} and crops \citep{Yang2012-aw} and exploited in livestock improvement efforts \citep{Mulder2008,Ibanez-Escriche2008-ie}.

@@

@@

We have demonstrated typical usage of the \texttt{R} package \texttt{vqtl} for mean-variance QTL mapping in an F2 intercross.
This package is appropriate for crosses and phenotypes where genetic factors or covariates or are known or suspected to influence phenotype variance.
In the case of genetic factors, they can be mapped, as illustrated in \CortyReanalysisPaper.
In the case of covariates, they can be accommodated, which can increase power and improve false positive rate control, as described in \CortyMethodsPaper.

>>>>

It has long been recognized that the extent of residual variation is itself a heritable trait \citep{Falconer1965,Lynch1998}.
This phenomenon has been described theoretically \citep{Hill2004-uo,Hill2010} and demonstrated in inbred model organisms \citep{Sorensen2015,Ayroles2015} and crops \citep{Yang2012-aw,Forsberg2015} and exploited in livestock improvement efforts \citep{Mulder2008,Ibanez-Escriche2008-ie}.

@@

In two companion articles, we apply mean-variance QTL mapping with package \texttt{vqtl} to discover new QTL.
In the first, we describe a novel vQTL for exploratory behavior and a novel mQTL for a circadian behavior trait, both in mouse F2 intercrosses retrieved from the Mouse Phenome Database (\CortyReanalysisPaper) \citep{Bogue2015}.
The dataset from one of these intercrosses is used to benchmark the performance of package \texttt{vqtl}, as is described in the ``Performance Benchmarks'' section.
In the second, we describe a unique strength of mean-variance QTL mapping, its ability to accommodate variance heterogeneity arising from background factors, such as sex, batch, and housing, and demonstrate the value of such accommodation with both simulation studies and a report of a novel QTL (\CortyMethodsPaper).

@@

We have demonstrated typical usage of the \texttt{R} package \texttt{vqtl} for mean-variance QTL mapping in an F2 intercross.
This package is appropriate for crosses and phenotypes where genetic factors or covariates or are known or suspected to influence phenotype variance.
In the case of genetic factors, they can be mapped, as illustrated in one companion article (\CortyReanalysisPaper).
In the case of covariates, they can be accommodated, which can increase power and improve false positive rate control, as illustrated in another companion article (\CortyMethodsPaper).
